title: "Data Analysis Project"
author: "Marijke Friedman, Brendan Weissel, Sade Ajishegiri"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggthemes)

```


```{r}
ev_raw <- read_csv("data/EV_RAW_DATA.csv") 

ev_raw <- ev_raw |>
  clean_names()
glimpse(ev_raw)
```

```{r}
#votes by day 
votes_by_day <- ev_raw |> 
  summarise(across(starts_with("day"), \(x) sum(x, na.rm = TRUE)))

#votes by party
votes_by_party <- ev_raw |>
  group_by(party_code) |> 
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE)) |> 
  arrange(party_code)

### votes by precinct 
votes_by_precinct <- ev_raw|>
  group_by(precinct)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(precinct)

### votes by gender
votes_by_gender <- ev_raw|>
  group_by(gender_code)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(gender_code)

### votes by age
votes_by_age <- ev_raw|>
  group_by(age_group)|>
  summarise(total_votes = sum(day1, day2, day3, day4, day5, day6, day7, day8, na.rm = TRUE))|>
  arrange(age_group)
```

### Our original questions:

1. What age group had the highest turn out for early voting?
2. What county had the highest turnout for early voting?
3. How does the 2024 early voting data compare to the 2020 early voting data?
4. What counties had the closest margin between registered Republicans and Democrats who voted early? Can we identify any swing counties from this?
5. How do early voting patterns vary among different demographic groups? 

###: Our new questions:

1. What precincts did younger age groups dominate voting early in? What do these precincts have in common?
2. Are there any precincts where certain demographic groups outperformed their Census data? For example, are there precincts were 65% of the early voters were women despite women only being 50% of the voting age population?
3.  Which precincts had the most early voters identifying outside of male/female? How did they vote?


```{r}
#Trying to answer Q3 - Sade

#comments from Derek: "Much better job of refining your questions in this update. I suspect that the relatively low number of early voters not identifying as either male or female might be relatively quick to answer, so maybe tackle that one first?"

#"To answer your question about the gender_code values, the Maryland voter registration form (https://elections.maryland.gov/voter_registration/documents/English_Internet_VRA.pdf) offers three choices: Male, Female, and "Unspecified or Other". That last one isn't super helpful in terms of precision, because there could be a big difference between declining to say and affirmatively identifying as non-binary, for example. But it's the best you have. I would be very interested in the age and party groupings of those voters, fwiw."

#Finding ages/parties of unspecified/other gender identifiers - - - 

```

```{r}
#how many gender codes are there AND what percentage of NB early voters were there?

ev_raw |> 
  group_by(gender_code) |> 
  count()

```

```{r} 
(2777/(80405+77290+2777))*100
  
```
1.7% of early voters were effectively non binary.

```{r}

#trying to find precincts with the most nonbinary voters:

nonbinary_voters <- ev_raw |> filter(is.na(gender_code))

nonbinary_voters<- nonbinary_voters|>mutate()



```


4. Which precincts have historically turned out more early voters than others? How did it compare this election?
5. Do most precincts follow the general pattern of having the most people vote on the first and last days of early voting? Which precincts differ from this pattern?

### How we've refined our questions:

Based on your feedback, we're refining our questions to focus more on precinct-level data instead of overall state data. We are looking more closely at patterns that stand out in certain precincts - taking note of the unusual and trying to find the reasons behind it.

### Describing any obvious limitations/flaws in our data - Why might we be unable to answer our previous questions? Describe the limits of the data in terms of the time frame, what is and is not included and any codes or other information we'll need to find definitions for, and what we've done to better define it. 

The data we have is just early voting data. It's interesting to analyze because early voters are likely to be more civic duty-minded. But the early voter data doesn't tell us that much about the overall electorate, so it could be helpful to bring in actual results data and more information about the overall electorate.
We'll likely use Census data so we can make more meaningful comparisons, plus voter registration data so we can look at voter turnout numbers. 
As far as defining some of the terms in our questions and within our data, in the votes_by_gender table, what does "NA" actually mean? Did voters in that category mark a box? Did they write in something other than male or female? Was there an option that said "prefer not to say"?
