---
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

title: "Class Reference" author: "Sade Ajishegiri" output:
html_document: theme: cerulean highlight: pygments toc: tru toc_float:
collapsed: true smooth_scroll: false

## Introduction

Consider this a personal guide to the commands and functions you will
learn. In general, when you come across an R command or function that
you want to remember, put it in here along with a description of what it
does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this
sets the working directory to a directory called
data_journalism_2024_fall inside a directory called jour472 located
inside your home directory (the \~ is equivalent of /Users/[youruser] on
a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to assign a value to a variable

"label\<-number"

```{r}
coolnumber <-5
```

### How to run code

ctrl + shift + enter

### How to add a new chunk

Go to top right corner of this window and click the lil green c button
with a plus, then select "R"
