---
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

title: "Class Reference" author: "Sade Ajishegiri" output:
html_document: theme: cerulean highlight: pygments toc: tru toc_float:
collapsed: true smooth_scroll: false

## Introduction

Consider this a personal guide to the commands and functions you will
learn. In general, when you come across an R command or function that
you want to remember, put it in here along with a description of what it
does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this
sets the working directory to a directory called
data_journalism_2024_fall inside a directory called jour472 located
inside your home directory (the \~ is equivalent of /Users/[youruser] on
a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse:

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to assign a value to a variable:

"label\<-number"

```{r}
coolnumber <-5
```

### How to run code:

ctrl + shift + enter

### How to add a new chunk:

Go to top right corner of this window and click the lil' green c button
with a plus, then select "R"

### \~ Lab 1 \~

-   "group by" and "summary" stuff, AGGREGATES, means, medians, etc.

-   "How much" = total; sum "How many" = counting (n) function

-   DO NOT use the `readRDS` function but use `read_rds()`

### To store something as a variable object:

whatever you wanna name the object \<- code

### Questions:

-   what's the diff between .rds and .rmd files?

### Prelab 2:

-   changing data into a more useful version of itself

-   working with dates

-   mutate

-   broom icon in top right gets rid of shit in the environment window

-   read_csv NOT read.csv; read_csv is a tidyverse function and does
    more

R sorts data by case, as well - it doesn't know that BALTIMORE and
Baltimore are the same city

takes a stringed value and uppercases it so that R reads them as the
same

`readr` anticipates some date formatting and can automatically handle
many of these issues (indeed it uses lubridate under the hood). When you
are importing a CSV file, be sure to use `read_csv`, not `read.csv`.

DIFFERENT WAYS TO LOOK AT DATASETS:

-   head(primary_24) - shows first 6 rows
-   summary(primary_24) - shows quartiles
-   colnames(primary_24) - lists all the columns
-   glimpse(primary_24) - shows metadata of the dataset; shows how many
    rows/columns, etc.

### Mutate: - can add a new column to the data based on existing columns/data

Filter function - always used double equals signs (==)

Be careful using the select function - can accidentally get rid of some
columns

AND is restrictive and intersectional

**Filtering creates a subset of the data based on criteria**. All
records where the amount is greater than 150,000. All records that match
"College Park". Something like that. **Filtering works with rows -- when
we filter, we get fewer rows back than we start with.**

### Sanity Check:

-   clear environment
-   run all and check for errors
-   this makes sure that the code's been created in the specific
    notebook

### LAB 2: CENTRAL IDEAS:
- people who registered to vote after July 21st, esp.
HOW TO SUMMARIZE: WILL ALWAYS USE A COLUMN

The code below - group_by collects all the separate weeks, then summarize and count(n()) counts up each item and puts it into the new column we just made, which is "voters_per_week"
```{r}
md_voters_2024|>
  mutate(weeks=week(COUNTY_REGISTRATION_DATE))|>
group_by(weeks)|>summarize(voters_per_week=n())
   
```


### LAB 3

- distinct() gets rid of duplicates
- 