---
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

title: "Class Reference" author: "Sade Ajishegiri" output:
html_document: theme: cerulean highlight: pygments toc: tru toc_float:
collapsed: true smooth_scroll: false

## Introduction

Consider this a personal guide to the commands and functions you will
learn. In general, when you come across an R command or function that
you want to remember, put it in here along with a description of what it
does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this
sets the working directory to a directory called
data_journalism_2024_fall inside a directory called jour472 located
inside your home directory (the \~ is equivalent of /Users/[youruser] on
a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")
```

### How to install and load the tidyverse:

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### How to assign a value to a variable:

"label\<-number"

```{r}
coolnumber <-5
```

### How to run code:

ctrl + shift + enter

### How to add a new chunk:

Go to top right corner of this window and click the lil' green c button
with a plus, then select "R"

### \~ Lab 1 \~

-   "group by" and "summary" stuff, AGGREGATES, means, medians, etc.

-   "How much" = total; sum "How many" = counting (n) function

-   DO NOT use the `readRDS` function but use `read_rds()`

### To store something as a variable object:

whatever you wanna name the object \<- code

### Questions:

-   what's the diff between .rds and .rmd files?

### Prelab 2:

-   changing data into a more useful version of itself

-   working with dates

-   mutate

-   broom icon in top right gets rid of shit in the environment window

-   read_csv NOT read.csv; read_csv is a tidyverse function and does more

R sorts data by case, as well - it doesn't know that BALTIMORE and Baltimore

takes a stringed value and uppercases it so that R reads them as the same

DIFFERENT WAYS TO LOOK AT DATASETS:

-   head(primary_24) - shows first 6 rows
-   summary(primary_24) - shows quartiles
-   colnames(primary_24) - lists all the columns
-   glimpse(primary_24) - shows metadata of the dataset; shows how many
    rows/columns, etc.

### Mutate: - can add a new column to the data based on existing columns/data

Filter function - always used double equals signs (==)

Be careful using the select function - can accidently get rid of some columns

AND is restrictive and intersectional

### Sanity Check:

-   clear environment
-   run all and check for errors
-   this makes sure that the code's been created in the specific
    notebook
