---
title: "lab_05"
author: "derek willis"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# load the tidyverse
options(scipen=999)
library(tidyverse)
library(lubridate)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. This week, we'll revisit our new Maryland voters data and try to join it with contributions to the Harris campaign from Maryland donors made via ActBlue. Can we find new voters who are not registered Democrats who gave to Harris after she became a presidential candidate on July 21? What, if anything, can we say about those people?

We'll be doing something similar to what Ence Morse did for [this Washington Post story](https://www.washingtonpost.com/politics/2024/09/24/kamala-harris-younger-donors/?pwapi_token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJyZWFzb24iOiJnaWZ0IiwibmJmIjoxNzI3MzIzMjAwLCJpc3MiOiJzdWJzY3JpcHRpb25zIiwiZXhwIjoxNzI4NzA1NTk5LCJpYXQiOjE3MjczMjMyMDAsImp0aSI6ImIwYWMxNTk4LWMwMDgtNDE5Ny04MjAyLWVmMWNjZWFmYWZiMiIsInVybCI6Imh0dHBzOi8vd3d3Lndhc2hpbmd0b25wb3N0LmNvbS9wb2xpdGljcy8yMDI0LzA5LzI0L2thbWFsYS1oYXJyaXMteW91bmdlci1kb25vcnMvIn0.3-s0tBkMV7Y9DD21CBsKMlaliGlOKT2egCO-UFT735Y), just on a smaller scale.

First, load two CSV contribution files in the `data` folder, one each for donations made in July and August, and then combine them into a single dataframe.

```{r}
# load the Harris contribution data files and combine them

july_harris_contribs <- read_csv("data/july_harris.csv")
august_harris_contribs <- read_csv("data/august_harris.csv")

harris_combined_contribs <- bind_rows(list(july_harris_contribs, august_harris_contribs))
view(harris_combined_contribs)

```

Next, let's read in the new voters (people who registered in 2024):

```{r}
# load the new voters data
new_voters <- read_csv("data/new_voters.csv")
```

Next join the Harris donations to the voter data. Remember, people's names are common, __so you'll want to use name columns and address information__. You'll need to figure out which columns to use from both dataframes. The new_voters data has both a single Address column and various parts of the address, such as HouseNumber. Make a plan for picking your join columns - try out some combinations - and save the results to a new dataframe.


```{r}
# join the combined Harris file to the new voters
# you will need to determine what columns to use

harris_contributors_and_new_voters <- harris_combined_contribs |> left_join(new_voters, join_by(address_one==Address))
```
NOTE: Remember that using a left join takes EVERYTHING from the left, will SHOW me everything from the left, but then ONLY WHAT matches from the right. (Do I have that right?)


```{r}
join_by_zip <- harris_combined_contribs|> inner_join(new_voters, join_by(zip==ResidentialZip))
```
NOTE: Did I overload R?


```{r}
new_harris_voters <- new_voters|> inner_join(harris_combined_contribs, join_by(LastName==last_name, Address==address_one, ResidentialZip==zip))
```
```{r}
new_harris_voters <- new_voters|> inner_join(harris_combined_contribs, join_by(LastName==last_name,FirstName==first_name, Address==address_one, ResidentialZip==zip))
```
```{r}
new_harris_donors <- harris_combined_contribs|> left_join(new_harris_voters,join_by(last_name==LastName, first_name==FirstName, address_one==Address, zip==ResidentialZip))
```
wait wait ignore that idk what i was trying there

## Answer questions

Q1. Explain why you used the columns you did to make the join, and why that result is better than other ways you tried. Are there any problems with the join you made, such as the possibility of false positives?

A1. 

I used the first and last name, address and zip code columns to make the join because I think those are the most likely to be unique in combination. I think it's more likely that a number of our target demo slipped through the cracks than that we racked up a ton of false positives, but of course the possibility is still there. Some first/last name combos in the set are pretty common, and if two voters with the same name have a typo in their zip or address, it will probably result in a false positive.


Q2. Based on your join, how many donations to Harris have come from people who registered this year? How many unique people does that represent, and which donors have given the largest _number_ of donations since July 20? 

A2. 

Based on the join, I think at least 363 donations came from newly-registered people? Sophia Hoffer-Perkins, Brittany Musolino and Eric Snipe appear to be the donors who have given the most separate times at 7 donations each.

```{r}
new_harris_voters |> 
  group_by(LastName, FirstName, Address, ResidentialZip) |> 
  summarize(donations = n(), total = sum(amount)) |> 
  arrange(desc(donations))
```

Q3. What are the party and gender breakdowns of those donors? How many donations has Harris gotten from non-Democrats?

A3. 
All three of the donors mentioned above are democrats. Only 43 of her donations have been from non-democrats.

```{r}
new_harris_voters|>
  group_by(Party)|> summarise(donations = n(), total = sum(amount)) |> 
  arrange(donations)
```

Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve just the new voters, or the donors or the joined dataframe, but it cannot be a variation of a previous question (so, for example, you can't ask, "How many Republicans have given to Harris"). First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.

A4:
Question: Where do Harris' top 10 donors (people that have given her the largest donations) live?



```{r}

```

