---
title: "lab_04"
author: "Derek Willis"
date: "2024-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(janitor)
library(lubridate)

```

For this lab, we want to investigate spending by the two leading Senate candidates in Maryland during this election, Angela Alsobrooks and Larry Hogan. Our goal is to try and standardize the purpose of the expenditures so that we can compare the two campaigns. We have the data, but it's messy - purpose descriptions are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis.

## Load Data

You'll need to load one data set, a CSV file of campaign expenditures located in the data folder called "md_senate_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table
md_expenditures <- read_csv("data/md_senate_expenses.csv")

```

## Answer questions

**Q1.** You've been assigned to compare the spending priorities of the two campaigns, but they describe their spending differently using the `purpose` column. Using OpenRefine, create a project using the original CSV file, make a copy of the `purpose` column called `purpose_clean` and then standardize it, focusing on the purposes that appear most often or have the largest total amounts. You can transform the data as you want - making all records the same case, for example, to help. The goal is to be able to group together as many similar purpose values, so you can choose to call all transportation-related expenses the same thing.

How do you know when you're done? It's a judgment call, but you should be close to 100 unique values for `purpose_clean`, and maybe even below that number.

Then, when you are done standardizing `purpose_clean`, export it as a new CSV file to your data folder, *giving it a different name* and read it back into RStudio as a new dataframe.

Then, using that `purpose_clean` column, write code to find the largest amounts of spending on categories by both campaigns.

NOTE: "AMOUNTS" -  SO SUM!
```{r}
md_expenditures <- read_csv("data/lab_04_md_senate_expenses.csv")

```


**A1.**

Both Alsobrooks and Hogan spent the most on advertising.
```{r}
md_expenditures |>

    group_by(candidate,purpose_clean)|>
   summarize(most_spent = sum(amount)) |> 
  arrange(desc(most_spent))

  
```

NOTE: COUNTING VS SUMMING - counting is how many individual pieces of something; summing is what those EQUAL - like 5 pieces of paper money could sum to 5 dollars, but it could also sum up to 500 dollars.

**Q2.** Let's do some research on the top purposes from A1. Feel free to search the Web to understand those terms better if need be, and then write a description comparing and contrasting the two campaigns' largest categories of spending.

**A2.**
Alsobrooks spent most of her funds on ads, likely because Marylanders don't really know who she is yet, while Hogan, in contrast, has been governor twice. She also spent a good deal in the mailing lists category, which includes payments to third-party services that help candidates compile lists of voters and reach out to them. I'm guessing for a similar reason to the above.

Hogan did spend most of his funds on advertising,too, but he spent much less than Alsobrooks - again, probably because most voters in Maryland already know who he is. I think it's kind of ironic that one of his top spending categories is "fundraising services" - a label his campaign used. Researching some of the companies paid for this category, it looks like it's going to campaign fundraising consulting firms.

**Q3.** Pick one of the top categories for each campaign and write code that shows how that spending has changed over time, if at all. You'll need to use lubridate for this. Write a couple of sentences describing that pattern.

**A3.**
It looks like for both candidates, they don't spend as much on advertising at the very beginning of the campaign, which surprised me. Both Hogan and Alsobrooks appear to have spent the most on advertising around April, May and June.
```{r}
md_expenditures|>
  filter(candidate==Hogan)|>
  
mutate(by_month = floor_date(date,"month")) |>
 
   group_by(by_month) |>
   summarize(total_month = sum(amount)) |> 
  arrange(desc(total_month))
```
NOTE: THE ABOVE ERROR HAPPENED, I THINK, BECAUSE I DIDN'T PUT HOGAN IN PARENTHESES.

```{r}
md_expenditures|>
  filter(candidate=="Hogan")|>
  
mutate(by_month = floor_date(date,"month")) |>
 
   group_by(by_month) |>
   summarize(total_month = sum(amount)) |> 
  arrange(desc(total_month))
```
NOTE: OKAY BUT I NEED TO SEE TOP SPENDING IN A CATEGORY, NOT OVERALL.
```{r}
md_expenditures|>
  filter(candidate=="Hogan")|>
  filter(purpose_clean=="ADS")|>
mutate(by_month = floor_date(date,"month")) |>
 
   group_by(by_month) |>
   summarize(total_month_on_ads = sum(amount)) |> 
  arrange(desc(total_month_on_ads))
```
```{r}
md_expenditures|>
  filter(candidate=="Alsobrooks")|>
  filter(purpose_clean=="ADS")|>
mutate(by_month = floor_date(date,"month")) |>
 
   group_by(by_month) |>
   summarize(total_month_on_ads = sum(amount)) |> 
  arrange(desc(total_month_on_ads))
```


**Q4.** Which committee has the larger payroll or salary spending so far? How many unique recipients of payroll/staff spending are there for each candidate?
```{r}
md_expenditures|>
  filter(candidate=="Alsobrooks")|>
  filter(purpose_clean=="PAYROLL TAXES/FEES")|>
mutate(to_date = floor_date(date,"year")) |>
 
   group_by(to_date) |>
   summarize(payroll_spending = sum(amount)) |> 
  arrange(desc(payroll_spending))
```
```{r}
md_expenditures|>
  filter(candidate=="Alsobrooks")|>
  filter(purpose_clean=="SALARY")|>
mutate(to_date = floor_date(date,"year")) |>
 
   group_by(to_date) |>
   summarize(payroll_spending = sum(amount)) |> 
  arrange(desc(payroll_spending))
```
There's gotta be a way to do that all at once.
```{r}
md_expenditures|>
  filter(candidate=="Hogan")|>
  filter(purpose_clean=="PAYROLL TAXES/FEES")|>
mutate(to_date = floor_date(date,"year")) |>
 
   group_by(to_date) |>
   summarize(payroll_spending = sum(amount)) |> 
  arrange(desc(payroll_spending))
```

**A4.**
Alsobrook's campaign has spent the most on payroll so far. Is it correct that Hogan hasn't spent anything on payroll/salary stuff??

**Q5** What's the most interesting story idea you've found in this data?

**A5**
If I haven't made a grave error and Hogan has indeed not spent anything on payroll fees, taxes or salary fees, then I think that's pretty weird - especially compared to what Alsobrook's campaign has spent in that category. If he's filing it under a different name, what is that name and why?