---
title: "lab_09.Rmd"
author: "derek willis"
date: "2024-10-27"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Datawrapper account
-   Election data in CSV files in the `data` folder
-   Tidycensus

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings
for this lab.

```{r}
# load libraries here

library(tidyverse)
library(janitor)
```

## Questions

**Q1.** You're interested in seeing how the methods of voting have
changed in Maryland since 2018. Using the `md_votes_by_type.csv` that
has data from the top of the ticket (governor in 2018 and 2022,
president in 2020), make a chart showing the percentage of votes cast in
each category - early voting, election day and mail - over the three
years. You'll need to make those percentages in R and then export the
CSV using `write_csv`, then make the chart in Datawrapper. Have a title,
description and source that tells readers about the data. The title and
description should highlight the most newsworthy or important aspect.

HINT: In Datawrapper's "Check and Describe" tab, you can click on a
column to get the option of "hiding" it in any visualization.

What do you think the story here is? Put the answer below along with the
URL of your chart.

**A1.** <https://datawrapper.dwcdn.net/52nlU/1/>;

I think the story is in the rise of mail-in voting due to the pandemic.
I'm curious to see if it will taper off even more for future elections.

```{r}
#Percentages

read_csv("data/md_votes_by_type.csv")

md_votes_by_type<- md_votes_by_type|>
  mutate(Percentage_mail_in = (Mail/Total)*100, Percentage_election_day = (`Election Day`/Total)*100, Percentage_early = (Early/Total)*100)

md_votes_by_type

write_csv(md_votes_by_type, "data/md_votes_by_type.csv")


```

#Mail-ins 2018: (162814/2304512)*100 #Election Day 2018:
(1482029/2304512)*100 #Early 2018: (659669/2304512)\*100

#Mail-ins 2020: (1502986/3037030)*100 #Election Day 2020:
(437126/3037030)*100 #Early 2020: (983985/3037030)\*100

#Mail-ins 2022: (338748/1754475)*100 #Election Day 2022:
(1037176/1754475)*100 #Early 2022: (378551/1754475)\*100

**Q2.** You've been assigned to track early-voting turnout in Maryland
at the county level. To do that, use [data from the
state](https://elections.maryland.gov/press_room/2024_stats/PG24/EarlyVoting%20RAW%20data.csv).
YOU DO NOT NEED TO LOAD THIS FILE. I've created several CSV files from
it; each has a row for each day, with the counties as columns, so you
can track early voting turnout in certain categories. The available
files in the data folder are:

-   ev_totals_18_24.csv - early voting turnout by the 18-24 age group
-   ev_totals_65_up.csv - early voting turnout by the 65+ age group
-   ev_totals_rep.csv - early voting turnout by Republicans
-   ev_totals_dem.csv - early voting turnout by Democrats
-   ev_totals_una.csv - early voting turnout by unaffiliated voters
-   ev_totals_women.csv - early voting turnout by women
-   ev_totals_men.csv - early voting turnout by men

Which file you use to make your chart, and what kind of chart is up to
you (but it MUST be a chart, not a map or table), so experiment to see
what works best. Since we have data over time, you might start with a
line chart, but you might not end up there. You can display all
counties, or just some of them - to limit the number of counties, you'll
need to write code to select only the columns that you need and then use
`write_csv()` to export the dataframe to a CSV file that you'll use in
Datawrapper.

Your chart MUST:

\* Have a title, description and source that tells readers about the
data. The title and description should highlight the most newsworthy or
important aspect. \* Easily readable labels and values \* Be published
\* Put the question you chose to answer below along, with the URL of
your chart.

Tell me below why you picked the CSV you did. If you made any changes to the data, put that code below.

**A2.**
https://datawrapper.dwcdn.net/9Am9x/1/; I picked this CSV because it was easy and the democrats early voter pattern was a little freaky 

```{r}
library(readr)
ev_totals_rep <- read_csv("data/ev_totals_rep.csv")
View(ev_totals_rep)

ev_totals_rep <- ev_totals_rep|> select(Day, `Baltimore County`, Montgomery, `Prince George's`
  
)

ev_totals_dem <- read_csv("data/ev_totals_dem.csv")
View(ev_totals_dem)

ev_totals_dem <- ev_totals_dem|> select(Day, `Baltimore County`, Montgomery, `Prince George's`)
  
ev_totals_una <- read_csv("data/ev_totals_una.csv")
View(ev_totals_una)

ev_totals_una <- ev_totals_una|> select(Day, `Baltimore County`, Montgomery, `Prince George's`)
```

```{r}
#RENAMING 
ev_totals_dem <- ev_totals_dem|> rename(balt_county_dem=`Baltimore County`, montgomery_dem=Montgomery, pg_county_dem=`Prince George's`)
```
```{r}
ev_totals_una <- ev_totals_una|> rename(balt_county_una=`Baltimore County`, montgomery_una=Montgomery, pg_county_una=`Prince George's`)

ev_totals_rep <- ev_totals_rep|> rename(balt_county_rep=`Baltimore County`, montgomery_rep=Montgomery, pg_county_rep=`Prince George's`)
```

```{r}
#JOINING
rep_and_una<- ev_totals_rep|> inner_join(ev_totals_una)

all_joined<- rep_and_una|>inner_join(ev_totals_dem, join_by (Day))

all_joined
```
```{r}
#NOW WRITE IT OUT TO A CSV FILE
write_csv(all_joined, "data/all_joined.csv")
```

```{r}
#THAT ONE SUCKED WRITING OUT JUST DEMS NOW
write_csv(ev_totals_dem, "data/dem_early_votes.csv")
```
  

Okay, so what do I actually want to do , though? I want to make a chart
showing early voters across counties - let's do the three most populated
counties - Montgomery, PG and Baltimore County.


**Q3.** After reading the recent CNN story about fundraising called
"[How elderly dementia patients are unwittingly fueling political
campaigns](https://www.cnn.com/interactive/2024/10/politics/political-fundraising-elderly-election-invs-dg/),"
you decide to dig into the habits of Maryland political donors who give money via ActBlue and WinRed.

You want to show Maryland's political geography - where those donors
live - and the frequency of donations. To do that, you'll load in two CSV files: `md_winred_zip_totals.csv` and `md_actblue_zip_totals.csv`.

```{r}

read_csv("data/md_winred_zip_totals.csv")

read_csv("data/md_actblue_zip_totals.csv")
```


You also will need to get population by zip code for Maryland - we did a similar thing in a previous lab, but this time you want the total population as your variable. Use tidycensus to do that, and save that dataframe, then join it to each of your contribution dataframes - you will need to make sure that the zip codes you're joining on are the same
datatype.

In each dataframe, use the *number of donations* to calculate a per
capita rate of the number of donations per 1,000 population for each zip
code using `mutate()`. Calculating a rate works like this: it's the
thing you're counting divided by the population times the "per" number.

Then export each dataframe using `write_csv` and make two (2) choropleth
maps in Datawrapper. For each map, you will need to:

-   Search for Maryland zip codes (use the 2020 data) as the base map
-   Upload the CSV file
-   Change the tooltip to show both the rate of donations and the total
    amount (you can put them side by side or put one under the other, as
    we did in the pre_lab). Make sure to properly format both numbers to
    make them readable.
-   Have a title, description and source that tells readers about the
    data (the source is the Federal Election Commission, and the data
    covers Jan. 1, 2023 through Oct. 16, 2024). The title and
    description should highlight the most newsworthy or important
    aspect.

Compare the two maps. What do you think are the most interesting story
ideas in both, and what would you want to know more about? Put the
answer below along with the URLs of your maps.

**A3.**

```{r}

```
